---
title: "Project"
author: "SeanJ- Volcaetus"
date: "4/3/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data  ##
From TidyTuesday
URL:https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-07-07


```{r }
coffee_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')

```
# Quick overview #

```{r}
summary(coffee_ratings)
```
>A few NA's.

>1 within quakers, and 230 in Altitude low/high/mean

> Check what is happening in the rest of the data set

# Count of NA's per coloumn#
```{r}
apply(X=is.na(coffee_ratings), MARGIN = 2, FUN = sum)
```
> I will be just removing some of the columns with many missing vlaues, for instance farm_name.

```{r}
library(tidyverse)
```

Removal of columns
```{r}
coffee = coffee_ratings%>%
  select(-farm_name,-lot_number,-mill,-ico_number,-altitude,
         -altitude_low_meters,-altitude_high_meters,-producer,-company,
         -expiration,-certification_address,-owner_1,-grading_date)
apply(X=is.na(coffee), MARGIN = 2, FUN = sum)
view(coffee)
```


```{r}
try = na.omit(coffee)
```










```{r}
#df only na's
NAs=coffee_ratings%>%
  select(everything())%>%
  filter(is.na(quakers | altitude_mean_meters | altitude_high_meters | altitude_low_meters))
#take a look at missing values within rows
NAs

```

```{r}
#df with no na's
coffee=coffee_ratings%>%
  select(everything())%>%
  filter((!is.na(quakers | altitude_mean_meters | altitude_high_meters | altitude_low_meters)))
```

```{r}
old = count(coffee_ratings)
new = count(coffee)
removed = old - new
pct.Kept = new / old  
pct.Removed = removed / old 
print(c(old,new,removed,pct.Kept,pct.Removed))
```
I would prefer to have less removed, but this captures at least 80% of the coffee rating data.
If the altitude does not show to have an effect on the coffee rating, then the removal of the altitude columns from the  original data will be done. Then re-running analyis on this set will occur at that time.


```{r}
coffee
```



